Phylogenetics

### 2024-04-04 episodes/msa.Rmd
Fixed a few things on MSA post-first edition Fixed a figure link on phylo

---
title: 'Multiple sequence alignment'
teaching: 0
exercises: 240
---

@@

```bash
$ interactive -A uppmax2024-2-10 -M snowy -t 4:00:00
$ cd /proj/g2020004/nobackup/3MK013/<username>
$ mkdir phylogenetics
$ cd phylogenetics
```

@@

```bash
$ cp ../../data/rpoB.fasta .
```

@@

```bash
$ grep '>' rpoB.fasta | head -n 5
```

@@

```bash
$ sed "/^>/ s/>\([^ ]*\) \([^\[]*\)\[\(.*\)\]/>\\3_\\1/"  rpoB.fasta | sed "s/ /_/g" > rpoB_renamed.fasta
$ grep '>' rpoB_renamed.fasta | head -n 5
```

@@

```bash
$ module load bioinfo-tools MAFFT
$ time mafft-fftns rpoB_renamed.fasta > rpoB.fftns.aln
```

@@

@@

```bash
$ time mafft-einsi rpoB_renamed.fasta > rpoB.einsi.aln
```

@@

@@

```bash
clipkit -h
```

>>>>

---
title: 'Multiple sequence alignment'
teaching: 0
exercises: 120
---

@@

```bash
interactive -A uppmax2024-2-10 -M snowy -t 4:00:00
cd /proj/g2020004/nobackup/3MK013/<username>
mkdir phylogenetics
cd phylogenetics
```

@@

```bash
cp ../../data/rpoB/rpoB.fasta .
```

@@

```bash
grep '>' rpoB.fasta | head -n 5
```

@@

```bash
sed "/^>/ s/>\([^ ]*\) \([^\[]*\)\[\(.*\)\]/>\\3_\\1/"  rpoB.fasta | sed "s/ /_/g" > rpoB_renamed.fasta
grep '>' rpoB_renamed.fasta | head -n 5
```

@@

```bash
module load bioinfo-tools MAFFT
time mafft-fftns rpoB_renamed.fasta > rpoB.fftns.aln
```

@@

Take a few minutes to read upon the different alignment strategies on the page above.

@@

```bash
time mafft-einsi rpoB_renamed.fasta > rpoB.einsi.aln
```

@@

As before, if you do not have access to a terminal on your windows laptop, use MobaXterm and Session > SFTP to copy files to your computer.

@@

```bash
module load ClipKIT
clipkit -h
```

### 2024-04-04 episodes/phylogenetics.Rmd
Fixed a few things on MSA post-first edition Fixed a figure link on phylo

The exercise is done for a large part with pen and paper, and then a demonstration in R on your laptop. You can already RStudio. We'll also use the R package [`ape`](https://emmanuelparadis.github.io/) , which you should install if it's not present on your setup:

```r
install.packages('ape')
```

@@

![UPGMA tree](fig/upgma_manual.png){alt='Manually built UPGMA tree'}

@@

## Exercise 2: maximum-likelihood tree

>>>>

The exercise is done for a large part with pen and paper, and then a demonstration in R on your laptop, using RStudio. We'll also use the R package [`ape`](https://emmanuelparadis.github.io/), which you should install if it's not present on your setup. Commands can be typed or pasted in the "Console" part of RStudio.

```{r}
install.packages('ape')
```

@@

![UPGMA tree](episodes/fig/upgma_manual.png){alt='Manually built UPGMA tree'}

@@

## Exercise 2: Neighbor-joining and Maximum-likelihood tree
